### 2017-12-01 2/tfpie/response.txt

The paper is well-written, is well-positioned in the context of the
status quo and balances the pedagogical approach with the material
used. The idea of language levels is properly motivated and the
continuation of this approach beyond the first semester seems - after
reading the paper - the Obvious Choice. What is lacking, however, is a
motivation for the point at which the transition to Java is made. Why
aren't e.g. iteration, mutability & typing introduced in the similar
vein as objects, class-hierarchies, etc.? Was this a choice from
practicality, or is there a reason to introduce them post-transition?

RESPONSE: We clarify why the transition point was chosen.

 - (Sec.1) "interaction-oriented language" / "batch-oriented language"
   --> it seems the authors are refering to the environments/tooling
   rather than the language. If not, this terminology warrents
   clarification.

RESPONSE: That Java has no semantics for single-expression evaluation
is a fact about the language, not its tools.

@@

 - (Sec.2) "while most students do some software development, there is
   almost no uniformity beyond that" & "in which they will use
   professional-grade languages and tools which will almost certainly
   be object-oriented" are, at least, paradoxical (if not
   contradictory). This conflict in the motivation should be addressed
   for the final version of the paper.

RESPONSE: We edited this sentence.

- (Sec.6) "The class has also had notable success in the recruitment
  and retention of female students, as compared to the other versions
  of the second-semester course." --> I find this a rather hefty
  statement. It would - in my mind - at least warrent further
  research. If this consistently holds true, there are valuable
  insights to be gained about pedagogical gender-script in the field
  of CS. The statement by itself seems a bit hefty to be included
  without at least some concrete data and/or qualitative evaluation.

RESPONSE: We agree that this merits further study.  However, our
remark is anecdotal, and we have clarified it.

### 2017-12-01 2/admin/charter/CS3500_Report.tex
Move Fundamentals II content into own dir.

\title{CS~3500 Course Charter \& Syllabus}

\author{Jay Aslam, Will Clinger, David Van Horn \\
        College of Computer and Information Science \\
        Northeastern University}

\date{February 21, 2012}

@@

\section{CS~3500 Course Charter}

@@

\paragraph{Proposed Course Description:}
%
Presents the theory and practice of object-oriented programming,
emphasizing abstractions and ideas that scale to large systems.
Introduces representation independence and related concepts such as
information hiding, encapsulation, modularity, and interchangeability
of parts.  Analyzes strengths and weaknesses of alternative
representations by considering asymptotic efficiency for the best,
worst, average, and amortized case.  Illustrates general paradigms
such as divide-and-conquer algorithms, precomputation, caching,
memoization, and dynamic programming.  Covers static types and
explains how subtyping and several forms of polymorphism
(parametric, ad hoc, and inclusion) support design patterns that
contribute to reuse and extensibility.

@@

\section{CS~3500 Sample Syllabus}

>>>>

\title{Report from the CS3500 Committee}

\author{Jay Aslam, Will Clinger, David Van Horn\\
        College of Computer and Information Science \\
        Northeastern University}

\date{DRAFT: May 27, 2012}

@@

The committee was charged with reviewing the CS3500 (\emph{Object-oriented Design}) curriculum, with the end goal of revising CS3500 in such a way that
%
\begin{itemize}
\item it retains the best aspects of the current object-oriented design course,
\item it does not excessively and unnecessarily overlap with CS2510 (\emph{Fundamentals~2}), and
\item it contains substantially more algorithmic content.
\end{itemize}
%
Specific concerns included:
%
\begin{enumerate}
\item That CS2510 is teaching substantial amounts of object-oriented design.  Therefore, there is a concern that there is currently a fair amount of overlap between CS2510 and CS3500.
\item That students have reported that in co-op interviews they are asked about algorithms and data structures, but do not know how to answer these questions.
\item That CS3500 is serving, in part, as a review and consolidation course, due to the variability in how CS2510 is taught: Students entering CS3500 have mixed backgrounds in various programming topics, depending on the specific instance of CS2510 that they took, necessitating that CS3500 spend a fair amount of time bringing everyone up to speed.
\end{enumerate}

Given the above charge and specific concerns, the committee chose to review the three course introductory programming sequence (CS~2500 \emph{Fundamentals~1}, CS2510, and CS3500) in its entirety for content, redundancy, and the possible inclusion of algorithmic concepts.

Our committee began by (1) studying the extant course charters and sample syllabi for the three course programming sequence, noting differences with the courses as currently taught, (2) identifying redundancies and inconsistencies in the programming sequence as currently taught, and (3) identifying those object-oriented programming concepts that could reasonably be moved from CS3500 to CS2510.

The end result of eliminating inconsistencies and excessive redundancies, as well as moving a handful of programming concepts from CS3500 to CS2510 where natural and practicable, was to free up approximately three weeks of time within CS3500 that could be devoted to algorithmic concepts.  Upon conferring with Deans Finkelstein and Rasala, it was decided that this would be sufficient for the time being.

Two options were considered for adding algorithmic concepts to CS3500: (1) devoting a three week block to algorithms or (2) integrating algorithmic concepts throughout the course.  It was decided that, pedagogically, Option~(2) was preferable to Option~(1).  CS3500, as currently taught, already covers a number of topics in algorithms such as big-O notation and some advanced data structures (e.g., red-black trees), and there are natural opportunities to introduce other data structures and algorithmic concepts such as the divide-and-conquer, greedy, and dynamic programming paradigms.  These additional algorithmic concepts should be introduced and taught, where appropriate, within the course.

In what follows, we provide updated course charters and sample syllabi for CS2500, CS2510, and CS3500, reflecting our recommendations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% CS2500
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagebreak

\section*{CS~2500 Course Charter}

\paragraph{Course Title:}
%
CS~2500 \emph{Fundamentals of Computer Science 1} (4 SH)

\paragraph{Course Description:}
%
Introduces the fundamentals of program design via study of increasingly
complex data definitions: structures, containment, unions, self-referential,
mutually referential. In addition, the course studies design via generative
recursion, accumulator-style functions, and mutation of variables and
structures. The course also introduces models of computations, which explain
what programs do when they are applied to inputs.

\paragraph{Prerequisites:}
%
The course requires no previous experience in writing computer programs, and
only assumes mathematics at the level of high-school Algebra. Students are
expected to have an interest in learning about Computer Science in depth.

\paragraph{Textbooks:}
%
\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, 
MIT Press, 2000.

\paragraph{Topics Covered:}
%
Computers and the nature of computation. (Formal) models of computation.
Functions; function definitions. Conditional computation structures. Structure
definitions. Modeling information with data. The structure of data:
composite, containment, union, self-referential, mutually referential,
cyclically defined. Program design and design recipes: design based on the
structure of data. Extensional equality of structured data. (Automated) testing.
Accumulators. Designing abstractions. Generative recursion.

\paragraph{Course Outcomes:}
%
Upon completion of this course, a student should be able to \emph{design programs},
and, consequently, have some understanding of the fundamentals of computation.
In particular, the student should be able to: Complete data analysis and
definition for problems that require the use of composition, containment,
union, and self-reference. Design programs to solve problems that consume and
produce data structures based on composition, containment, union, and
self-reference. Design test suites for these programs and evaluate the
outcomes. Design simple abstractions and test the design. Reason about the
performance characteristics of programs in a basic way.

\paragraph{Measurement of Course Outcomes:}
%
Outcomes are measured and verified through: Weekly programming homeworks.
Daily short paper-and-pencil quizzes. Two to four major exams. Inclusion of
student's work in an electronic portfolio.

\pagebreak

\section*{CS~2500 Sample Syllabus}

The course sequences through the following topics:
\begin{itemize}
\item overview, intro to simple arithmetic calculation
\item atomic \& compound data: 
  \begin{itemize}
    \item integers
    \item booleans \& conditional computation
    \item records
  \end{itemize}
\item representation: information vs. data
\item data-driven programming
\item testing
\item unions
\item self-referential unions 1---recursion
\item self-referential unions 2---list processing
\item designing larger programs: top-down design; iterative refinement
\item recursion in all forms: trees, s-expressions, forests
\item abstraction: interfaces and implementations
\item abstraction: parametric data definitions and contracts
\item notions of equality
\item local name scope; simple asymptotic complexity
\item abstracting control with higher-order functions
\item constructing higher-order functions with lambda
\item infinite data structures and computations
\item generative recursion: root finding, sorting
\item generative recursion: graphs \& search; accumulators
\item review and wrap up
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% CS2510
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagebreak

\section*{CS~2510 Course Charter}

\paragraph{Course Title:}
%
CS~2510 \emph{Fundamentals of Computer Science 2}	(4 SH)

\paragraph{Course Description:}
%

Introduces the fundamentals of the design of class hierarchies in
object-oriented programming with emphasis on representation-dependent
program design with objects. The course also covers basic reasoning
techniques concerning the correctness and algorithmic behavior of
computations.

\paragraph{Prerequisites:}
%
CS~2500. In addition, CS~1800 Discrete Structures for Computer Science should be taken prior to CS~2510 or concurrently.

\paragraph{Textbooks:}
%
\emph{How to Design Programs}, Felleisen, Flatt, Findler, and Krishnamurthi, MIT Press, 2000. \emph{How to Design Classes}, Felleisen, Flatt, Findler, Krishnamurthi, Gray, and Proulx, in preparation.

\paragraph{Topics Covered:}
%
Objects, classes, class hierarchies.  Data definitions and methods.
%
Class diagrams.
%
Interfaces, abstract classes, subclasses.  Types.  Polymorphism.
%
Accumulators, iterators, loops.
%
Function objects, comparators, visitors.  Sorting: quick sort, merge
sort, heap sort.  Notions of equality.
%
Java collections library.
%
Overriding {\tt equals} and {\tt hashCode}.
%
Testing: unit testing, stress testing, and property-based random testing.
%
Binary trees, binary search, heaps, hash tables.
%
Invariants.
%
Graphs, graph traversal, generative recursion.  Depth-first,
breadth-first, and priority-queue based search.
%
Design patterns.
%
Mutation and stateful programming.
%
Indexed data structures.

\paragraph{Course Outcomes:}
%
Upon completion of this course, a student should be able to: 
(1) Design, implement, and test class hierarchies with appropriate methods. 
(2) Read the documentation of class libraries and write programs that use them appropriately. 
(3) Understand the algorithmic behavior of individual methods.

\paragraph{Measurement of Course Outcomes:}
%
Weekly programming homeworks. Final project. Weekly short paper and pencil quizzes. Two to four major exams. Inclusion of student's work in an electronic portfolio.

\pagebreak

\section*{CS~2510 Sample Syllabus}

The course studies the class-based program design and the design of abstractions that support the design of reusable software and libraries. It covers the principles of object oriented program design, the basic rules of program evaluation, the use of abstractions in support of the design of reusable code, and examines the relationship between algorithms and data structures, as well as basic techniques for analyzing algorithm complexity.

The course is suitable for both CS majors and non-majors. It assumes that student has been introduced to the basic principles of program design and computation.

\paragraph{Course Objectives:}
%
The goal is to help the student understand the principles of class
based program design using an object-oriented programming
language. Java is used to illustrate how the program design principles
are used in practical applications. Using a concrete language also
presents an opportunity to discuss the strengths and weaknesses of
languages and language paradigms.

\paragraph{MODULE 1:}
%
Review of Fundies 1 in the context of a class-based statically typed language (2 weeks)
%
\begin{itemize}
\item Review of the program design and abstractions in functional style;
  design recipe, abstractions, loops with accumulators
\item Defining classes and class hierarchies to represent complex information:
  simple classes, classes with fields that are instances of another class, 
  unions of classes, self-referential and mutually referential class   
  hierarchies: lists, trees, graphs
\item Designing methods for classes: static type system, dynamic dispatch,
  delegation (chain of responsibility)
\item Design Recipe for data definitions, for method definitions
\end{itemize}

\paragraph{MODULE 2:}
%
Designing Abstractions (3 weeks)
%
\begin{itemize}
\item Review of the Design Recipe for Abstractions
\item Abstracting over the data: abstract classes; subclasses
\item Abstracting over the functional behavior: function objects, functional iterator
\item Understanding equality of complex data
\end{itemize}

\paragraph{MODULE 3:}
%
Stateful Programming (3 weeks)
%
\begin{itemize}
\item Designing methods with effects; Designing tests for effects
\item Understanding equality in the presence of mutation
\item Direct access data structures: (e.g. Java ArrayList)
\item Loops
\end{itemize}

\paragraph{MODULE 4:}
%
Abstracting over the Data Type; Designing and using libraries (6 weeks)
%
\begin{itemize}
\item Polymorphism: parametrized classes, methods
\item Abstract Data Type --- stack, queue, list, set
\item Abstract Data Type --- map, hashmap
\item HashCode method and overriding equals method
\item Algorithm complexity --- estimates; Stress tests
\item Java Collection Framework
\end{itemize}

\paragraph{Notes:}
%
\begin{itemize}

\item A number of algorithms (filter, andmap, ormap, linear search, binary search, binary search trees, insertion sort, selection sort, quicksort, heap-based priority queue, heapsort, table lookup, use of hashmap, and some graph traversal algorithms (BFS, DFS, and possibly shortest path) are used to illustrate the program design, the need for appropriate data representation, and the need for understanding the complexity of algorithms.

\item Stress tests are used to reinforce the need for understanding the algorithm complexity.

\item Data structures (stacks, queues, lists, trees, arraylist, hashmap, lookup tables) are presented in the context of problems that motivate their use.

\item The goal is to make sure the students begin to understand the design of class based libraries, know how to look up the needed information, and can begin using libraries in their programs.

\item A comprehensive final project challenges the students to apply the design principles and the knowledge of the libraries to a creative larger program with up to 1000 LOC.

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% CS3500
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagebreak

\section*{CS~3500 Course Charter}

@@

\paragraph{Course Description:}
%
Presents the theory and practice of object-oriented programming,
emphasizing abstractions and ideas that scale to large systems.
Introduces representation independence and related concepts such as
information hiding, encapsulation, modularity, and interchangeability
of parts.  Analyzes strengths and weaknesses of alternative
representations by considering asymptotic efficiency for the best,
worst, average, and amortized case.  Illustrates general paradigms
such as divide-and-conquer algorithms, precomputation, caching,
memoization, and dynamic programming.  Covers static types and
explains how subtyping and several forms of polymorphism
(parametric, ad hoc, and inclusion) support design patterns that
contribute to reuse and extensibility.

@@

\pagebreak

\paragraph{For reference\ldots}

\pagebreak

\section*{CS~3500 Sample Syllabus}

### 2017-12-01 2/exams/exam4/2.tex
Move Fundamentals II content into own dir.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{22}
\
\begin{enumerate}
\item Design a \verb|TreeVisitor| interface for 2/3-trees, and implement the
\verb|visit| method in all of your classes from the previous problem.
Make sure you design the interfaces and methods so that a visitor can
return any type of data.

@@

\newpage
\item Assuming {\tt same} and {\tt equals} work as expected, is
the given {\tt hashCode} method valid?  That is, does it satisfy or
violate the law of {\tt hashCode}?  If it is valid, give an argument
for \emph{why}.  If it is not, give a counter-example to the law of
{\tt hashCode}.
\end{enumerate}
\end{problem}
\newpage

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem}\pts{22}
\
\begin{enumerate}
\item Design a \verb|TreeVisitor| interface for 2/3-trees, and implement the
\verb|accept| method in all of your classes from the previous problem.
Make sure you design the interfaces and methods so that a visitor can
return any type of data.

@@

\begin{solution}
\begin{verbatim}
public boolean same(UnPair<X> p) {
  return (this.left.equals(p.left) &&
          this.right.equals(p.right))
      || (this.left.equals(p.right) &&
          this.right.equals(p.left));
}

{5,7}.equals({5,7}) --> true
{5,7}.equals({7,5}) --> true
{5,5}.equals({5,7}) --> false
\end{verbatim}
\end{solution}

\newpage
\item Assuming {\tt same} and {\tt equals} work as expected, is
the given {\tt hashCode} method valid?  That is, does it satisfy or
violate the law of {\tt hashCode}?  If it is valid, give an argument
for \emph{why}.  If it is not, give a counter-example to the law of
{\tt hashCode}.

\begin{solution}
\begin{verbatim}
It is not valid.

Counterexample:
{5,7}.equals({7,5}) but 5.hashCode() does not equal
7.hashCode()
\end{verbatim}
\end{solution}
\end{enumerate}
\end{problem}
\newpage

@@

\begin{solution}
\begin{verbatim}
interface List<X> {
    void setTail(List<X> ls);
    void setTailAcc(List<X> ls, Cons<X> last);
}

class MT<X> implements List<X> {
    public void setTail(List<X> ls) {
        throw new RuntimeException("setTail on MT");
    }
    public void setTailAcc(List<X> ls, Cons<X> last) {
        last.rest = ls;
    }
}

class Cons<X> implements List<X> {
    X first;
    List<X> rest;
    Cons(X first, List<X> rest) {
        this.first = first;
        this.rest = rest;
    }
    public void setTail(List<X> ls) {
        this.rest.setTailAcc(ls, this);
    }
    public void setTailAcc(List<X> ls, Cons<X> last) {
        this.rest.setTailAcc(ls, this);
    }
}

// Plus tests
\end{verbatim}
\end{solution}

### 2017-12-01 2/web/12.su.reg/exams/exam1a.tex
Move Fundamentals II content into own dir.

%% -----------------------------------------------------------------------------
%% 
\pts{34}
\begin{problem}  Tonight is another meeting of the ACM, and they
  always serve pizza. So, we want to know a bit about what they serve.

Here is a class diagram that describes a pizza:

\begin{verbatim}
                    +--------------+
                    |    IPizza    |<-------------------+
                    +--------------+                    |
                     / \   / \   / \                    |
                     ---   ---   ---                    |
                      |     |     |                     |
        +-------------+     |     +-----------+         |
        |                   |                 |         |
        |                   |                 |         |
+-------------+      +-------------+   +-------------+  |
| Crust       |<--+  | Cheese      |   | Topping     |  |
+-------------+   |  +-------------+   +-------------+  |
| String name |   |  | String kind |   | IPizza base |--+ 
+-------------+   +--| Crust crust |   +-------------+
                     +-------------+
\end{verbatim}

(It may seem strange, but the toppings have no names.)

@@

 Make examples of data (i.e., instances) for this class
    hierarchy. Include an example of a thin crust pizza with mozarella
    cheese, and three toppings.

\begsol{\vspace{0.5in}}
[POINTS 3: 1 point for the \texttt{Crust} example, 1 point for the a 
\texttt{Cheese}, 1 point for the the specified example of
\texttt{Topping}]

@@

Design a method, \code{weight}, that computes the weight of the pizza
represented by a \code{IPizza}.

      The weight of \code{crust} is always \code{5}.  The weight 
      of a \code{Cheese} is \code{3}. Each
      topping contributes one to the total size.

@@

Design the method \code{replaceCheese} that produces an
    \code{IPizza} with the same toppings as this one,
    having the same crust but with the cheese replaced by the
    cheese of the given kind.

@@

Design the method \code{sameCheeseCrust} that checks whether this pizza
has the crust with the given name and the cheese of the given kind.

@@

Show the resulting templates for the \code{Cheese} class.

>>>>

%% -----------------------------------------------------------------------------
%% 
\pts{34}
\begin{problem}  Valentine's day is coming, so here is a nice gift you
  can give or receive. Of course, it needs to be in a nice box, and
  wrapped carefully.

Here is a class diagram that describes this gift:

\begin{verbatim}
                     +--------------+
                     |     IGift    |<------------------+
                     +--------------+                   |
                     / \   / \   / \                    |
                     ---   ---   ---                    |
                      |     |     |                     |
        +-------------+     |     +-----------+         |
        |                   |                 |         |
        |                   |                 |         |
 +-------------+      +--------------+  +------------+  |
 | Jewel       |<--+  | Box          |  | Wrap       |  |
 +-------------+   |  +--------------+  +------------+  |
 | String name |   |  | String kind  |  | IGift gift |--+ 
 +-------------+   +--| Jewel inside |  +------------+
                      +--------------+ 
\end{verbatim}

@@

 Make examples of data (i.e., instances) for this class
    hierarchy. Include an example of a ruby in a red box with three
    wrappings around it.

\begsol{\vspace{0.5in}}
[POINTS 3: 1 point for the \texttt{Jewel} example, 1 point for the a 
\texttt{Box}, 1 point for the the specified example of \texttt{Wrap}]

@@

Design a method, \code{size}, that computes the size of the gift
represented by a \code{IGift}.

      The size of \code{Jewel} is always \code{5}.  The size
      of a \code{Box} is \code{8} regardless of what it contains. Each
      wrap contributes one to the total size.

@@

Design the method \code{replaceBox} that produces an
    \code{IGift} wrapped in the same wrappings as this one,
    containing the same jewel  inside but with the box replaced by the
    box of the given kind.

@@

Design the method \code{sameJewelBox} that checks whether this gift
contains the jewel with the given name and is inside of the box of the
given kind.

@@

Show the resulting templates for the \code{Box} class.

### 2017-12-01 2/web/12.su.reg/exams/exam4.su.12.tex
Move Fundamentals II content into own dir.

%% -----------------------------------------------------------------------------
%% 
\pts{45}
\begin{problem}
An association list is a way of representing an association between
values.  For example, you might you use an association list to
associate students with their grades, or battleships with their
position, or bank account numbers with their account balances.

>>>>

%% -----------------------------------------------------------------------------
%% 
\pts{45}
\begin{problem}
An association list is a way of representing an association between
values.  For example, you might use an association list to
associate students with their grades, or battleships with their
position, or bank account numbers with their account balances.

### 2017-12-01 2/web/12.su.reg/exams/exam5.su.12.tex
Move Fundamentals II content into own dir.

\vfill
\centerline{\Large CS 2510 Exam 5 (Alternate) -- Summer 2012}

@@

%% -----------------------------------------------------------------------------
%% 
\pts{45}
\begin{problem}
Develop an implementation of {\tt Comparator<Square>} that orders
squares by their area in such a way that the smaller the area, the
\emph{greater} the square is considered.  So for example, there is no
square greater than the one with width $= 0$ and height $= 0$.

You may rely on the following definition of {\tt Square}:

\begin{verbatim}
// Represents a square with given height and width.
class Square {
  Integer width;
  Integer height;
  Square(Integer width, Integer height) {
    this.width = width;
    this.height = height;
  }
}
\end{verbatim}
\end{problem}

\newpage
\begin{problem}
Design a method:
\begin{center}
{\tt <T> Boolean noSameNeighbors(ArrayList<T> ls, Comparator<T> c)}
\end{center}
that determines if the given array list has no adjacent elements
(i.e.~neighboring elements) that are considered of equal size
according to the given comparator.
\end{problem}
\newpage
\begin{problem}
Design a method:
\begin{verbatim}
<A> Comparator<ArrayList<A>> dict(Comparator<A> ca)
\end{verbatim}
that consumes a comparator for {\tt A}s and and produces a comparator
for array lists of {\tt A}s that is the \emph{dictionary order} (aka
\emph{alphabetic order}).

In other words, given an ordering on individual elements, produce an
ordering on \emph{lists of} elements that corresponds to the ordering
used in dictionaries.  A dictionary uses an ordering on individual
letters (single elements) to make an ordering on strings of letters
(lists of elements) so that ``abc'' $<$ ``abcd'', ``abc'' $<$ ``adc'',
``abc'' $>$ ``ab'', and ``abc'' $=$ ``abc'', etc.
\end{problem}
\ 
\newpage
\ 
\newpage
\ 
\end{document}

>>>>

\vfill
\centerline{\Large CS 2510 Exam 5 -- Summer 2012}

@@

%% -----------------------------------------------------------------------------
%% 
\pts{45}
\begin{problem}
Develop an implementation of {\tt Comparator<Posn>} that orders
Cartesian-coordinates by their distance from the origin in such a way
that the further away from the origin, the smaller the position is
considered.  So for example, there is no position larger than (0,0).

You may rely on the following definition of {\tt Posn}:

\begin{verbatim}
// Represents (x,y) in Cartesian-coordinate system
class Posn {
  Integer x;
  Integer y;
  Posn(Integer x, Integer y) {
    this.x = x;
    this.y = y;
  }
}
\end{verbatim}
\end{problem}

\newpage
\begin{problem}
Design a method:
\begin{center}
{\tt <T> Boolean isSorted(ArrayList<T> ls, Comparator<T> c)}
\end{center}
that determines if the given array list is in ascending sorted
order according to the given comparator.
\end{problem}
\newpage
\begin{problem}
Here is a data definition for pairs:

\begin{verbatim}
// Represents a pair of A and B.
class Pair<A,B> {
  A left;
  B right;
  Pair(A left, B right) {
    this.left = left;
    this.right = right;
  }
}
\end{verbatim}
Design a method:
\begin{verbatim}
<A,B> Comparator<Pair<A,B>> lexi(Comparator<A> ca,
                                 Comparator<B> cb)
\end{verbatim}
that consumes a comparator for {\tt A}s and a comparator for {\tt B}s
and produces a comparator for pairs of {\tt A}s and {\tt B}s that is
the \emph{lexicographic order}.  Mathematically speaking, the
lexicographic order of a pair is defined as follows:
$$
(a,b) \leq (a',b')\text{ if and only if }a < a'\text{ or }(a = a'\text{ and }b \leq b')
$$
\end{problem}
\ 
\newpage
\ 
\newpage
\ 
\end{document}

### 2017-12-01 2/web/12.su.reg/exams/exam6.su.12.tex
Move Fundamentals II content into own dir.

\vfill
\centerline{\Large CS 2510 Exam 5 -- Summer 2012}

@@

%% -----------------------------------------------------------------------------
%% 
\pts{45}
\begin{problem}
Develop an implementation of {\tt Comparator<Posn>} that orders
Cartesian-coordinates by their distance from the origin in such a way
that the further away from the origin, the smaller the position is
considered.  So for example, there is no position larger than (0,0).

You may rely on the following definition of {\tt Posn}:

\begin{verbatim}
// Represents (x,y) in Cartesian-coordinate system
class Posn {
  Integer x;
  Integer y;
  Posn(Integer x, Integer y) {
    this.x = x;
    this.y = y;
  }
}
\end{verbatim}
\end{problem}

\newpage
\begin{problem}
Design a method:
\begin{center}
{\tt <T> Boolean isSorted(ArrayList<T> ls, Comparator<T> c)}
\end{center}
that determines if the given array list is in ascending sorted
order according to the given comparator.
\end{problem}
\newpage
\begin{problem}
Here is a data definition for pairs:

\begin{verbatim}
// Represents a pair of A and B.
class Pair<A,B> {
  A left;
  B right;
  Pair(A left, B right) {
    this.left = left;
    this.right = right;
  }
}
\end{verbatim}
Design a method:
\begin{verbatim}
<A,B> Comparator<Pair<A,B>> lexi(Comparator<A> ca,
                                 Comparator<B> cb)
\end{verbatim}
that consumes a comparator for {\tt A}s and a comparator for {\tt B}s
and produces a comparator for pairs of {\tt A}s and {\tt B}s that is
the \emph{lexicographic order}.  Mathematically speaking, the
lexicographic order of a pair is defined as follows:
$$
(a,b) \leq (a',b')\text{ if and only if }a < a'\text{ or }(a = a'\text{ and }b \leq b')
$$
\end{problem}
\ 
\newpage
\ 
\newpage
\ 
\end{document}

>>>>

\vfill
\centerline{\Large CS 2510 Exam 6 -- Summer 2012}

@@

%% -----------------------------------------------------------------------------
%% 
\begin{center}
A useful utility
\end{center}
You may assume, \textbf{for the purposes of examples only}, the existence of a
class {\tt Iter<X>} that implements {\tt Iterable<X>}.  It has a
somewhat magical constructor in that it can take \emph{any} number of
{\tt X}s as arguments.  Its {\tt iterator} method produces an iterator
that iterates through those arguments.

For example:
\begin{verbatim}
new Iter<Integer>().iterator().hasNext()  --> false
new Iter<Integer>(1).iterator().hasNext() --> true
new Iter<Integer>(1).iterator().next()    --> 1
Iterator<Integer> i =
  new Iter<Integer>(1,2,3).iterator();
i.next() --> 1
i.next() --> 2
i.next() --> 3
\end{verbatim}
\newpage
\pts{45}
\begin{problem}
Design a method
\begin{verbatim}
    Double avg(Iterable<Double> nums)
\end{verbatim}
that computes the average of a series of numbers.  It should raise an
exception if the series is empty.
\end{problem}
\newpage
\begin{problem}
Design a class
\begin{verbatim}
   class SumSquares implements Iterable<Integer>
\end{verbatim}
which is given an {\tt Iterable<Integer>} when constructed and whose
{\tt iterator} method produces an iterator that produces the sum of
the squares the given iterator produces.

For example, if given a {\tt new Nat()}, the iterable you wrote in lab
that produces the natural numbers {\tt 0, 1, 2, 3, 4}, \dots, then
iterator should produce {\tt 0, 1, 5, 14, 30}, etc.

\end{problem}
\newpage
\begin{problem}
Design a class
\begin{verbatim}
   class Zip<X,Y> implements Iterable<Pair<X,Y>>
\end{verbatim}
which is given an {\tt Iterable<X>} and an {\tt Iterable<Y>} when
constructed and whose {\tt iterator} method produces an iterator
that produces pairs of values: one from the {\tt X} series, one from
the {\tt Y} series.  The iterator should have no more elements
whenever either of the given iterators have no more elements.

You may rely on the following definition of {\tt Pair}:
\begin{verbatim}
   class Pair<X,Y> {
       X left;
       Y right;
       Pair(X left, Y right) {
           this.left = left;
           this.right = right;
       }
   }
\end{verbatim}

\end{problem}
\newpage
\ 
\newpage
\ 
\end{document}

### 2017-12-01 2/web/12.su.reg/exams/exam7.su.12.tex
Move Fundamentals II content into own dir.

\vfill
\centerline{\Large CS 2510 Exam 4 -- Summer 2012}

@@

%% -----------------------------------------------------------------------------
%% 
\pts{45}
\begin{problem}
An association list is a way of representing an association between
values.  For example, you might use an association list to
associate students with their grades, or battleships with their
position, or bank account numbers with their account balances.

The representation of an association list is much like a typical list,
except that every ``cons'' has \emph{two} elements and the rest of the
list.  One element is a ``key'', the other is its association.

Here is a data definition for association lists parameterized over the
types of keys and their values:

\begin{verbatim}
interface Assoc<K,V> {}

class EmptyAssoc<K,V> implements Assoc<K,V> {}

class ConsAssoc<K,V> implements Assoc<K,V> {
  K key;
  V val;
  Assoc<K,V> rest;
  ConsAssoc(K key, V val, Assoc<K,V> rest) {
    this.key = key;
    this.val = val;
    this.rest = rest;
  }
}
\end{verbatim}

\begin{enumerate}
\item Develop the visitor pattern for association lists.
\item Design an implementation of an association list visitor that
  looks up the (first) value associated to a given key.  If the key is
  not in the list, the visitor shoud throw an exception.
\item Using the double dispatch approach, design a {\tt same} method
  that determines if this association list is structurally equal to a
  given association list.
\end{enumerate}
\end{problem}
\newpage
\ 
\newpage
\ 
\end{document}

>>>>

\vfill
\centerline{\Large CS 2510 Exam 7 -- Summer 2012}

@@

%% -----------------------------------------------------------------------------
%% 
\pts{45}
\begin{problem}

A matryoshka doll, or a Russian nesting doll, refers to a set of
wooden dolls of decreasing size placed one inside the other.  The
innermost doll is usually a baby, and thus we can describe matryoshka
dolls as either a baby, or a layer with a matryoshka doll inside.

\begin{enumerate}
\item Design a data definition for matryoshka dolls.
\item Draw the class diagram corresponding to your data definition.
\item Develop the visitor pattern for matryoshka dolls.
\item Design a visitor for matryoshka dolls that computes the nesting
  depth of the doll.  The nesting depth of the baby is 0.
\item Have a great summer.
\end{enumerate}
\end{problem}
\newpage
\
\newpage
\
\end{document}
